Constraint Programming - Inside Gecode

### 2014-05-02 paper.tex
added example of non-overlapping rectangles

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{cite}
\usepackage{amsfonts}
\usepackage[pdftex]{graphicx}
\graphicspath{{../png/}}
\DeclareGraphicsExtensions{.png}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{algorithmic}
\usepackage{array}
\usepackage{mdwmath}
\usepackage{mdwtab}
\usepackage{eqparbox}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{stfloats}
\usepackage{url}
\hyphenation{op-tical net-works semi-conduc-tor}

@@

	Consider the problem
	\[SEND + MORE = MONEY\]
	where every character is a variable and the position in the word describes the weight.
	\begin{equation}
	\label{eq:linearConstraint}
	\begin{split}
		S \cdot 10^3 + E \cdot 10^2 + N \cdot 10^1 + D \cdot 10^0 + \\ 
		M \cdot 10^3 + O \cdot 10^2 + R \cdot 10^1 + E \cdot 10^0 = \\ 
		M \cdot 10^4 + O \cdot 10^3 + N \cdot 10^2 + E \cdot 10^1 + Y \cdot 10^0
	\end{split}
	\end{equation}

	To narrow the solution space a little bit down I state additional constraints:
	\begin{enumerate}
	\item All variables must have different values.
		\begin{equation}
			S \ne E \land S \ne N \land S \ne D \land \dots \land E \ne N \land E \ne D \dots
		\end{equation}	
	\item There must not be a trailing zero. Combined with the so-called all-different constraint \cite{allDifferent} from above this results in
		\begin{equation}
			S \ne 0 \land M \ne 0
		\end{equation}
	\end{enumerate}

@@

	\subsection{Constraint Types}
	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \ref{eq:linearConstraint}. Which statements are possible to make varies than very much from the programming language in which the problem is implemented and the used toolset. Typical constraints would be
	\begin{itemize}
		\item Equalities: $x = y + 2$, $x^2 + y^2 = 1$, \dots
		\item Inequalities: $x \ne y$, \dots
		\item Relations: $x < 2 y$, $e^x < 4$, \dots
	\end{itemize}

@@

	\begin{thebibliography}{1}
		\bibitem{handbookCP}
		F.~Rossi, P.van~Beek and T.Walsh, \emph{Handbook of Constraint Programming}, Amsterdam, The Netherlands: Elsevier, 2006, ISBN: 978-0-444-52726-4
		\bibitem{allDifferent}
		W.~van~Hoeve, \emph{The Alldifferent Constraint: A Survey}, http://www.andrew.cmu.edu/user/vanhoeve/papers/alldiff.pdf
		\bibitem{trendsInCP}
		F.~Benhamou, N.~Jussien, B.A.O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN: 978-1-905209-97-2
		\bibitem{gecode}
		http://www.gecode.org/
	\end{thebibliography}
\end{document}

>>>>

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{cite}
\usepackage{amsfonts}
\usepackage[pdftex]{graphicx}
\graphicspath{{../png/}}
\DeclareGraphicsExtensions{.png}
\usepackage[cmex10]{amsmath}
\interdisplaylinepenalty=2500
\usepackage{algorithmic}
\usepackage{array}
\usepackage{mdwmath}
\usepackage{mdwtab}
\usepackage{eqparbox}
\usepackage[caption=false,font=footnotesize]{subfig}
\usepackage{fixltx2e}
\usepackage{stfloats}
\usepackage{url}
\hyphenation{op-tical net-works semi-conduc-tor}
\usepackage{listings}
\usepackage{color}

\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{gray}{rgb}{0.5,0.5,0.5}
\definecolor{mauve}{rgb}{0.58,0,0.82}

\lstset{frame=tb,
  language=C++,
  aboveskip=3mm,
  belowskip=3mm,
  showstringspaces=false,
  columns=flexible,
  basicstyle={\small\ttfamily},
  numbers=none,
  numberstyle=\tiny\color{gray},
  keywordstyle=\color{blue},
  commentstyle=\color{dkgreen},
  stringstyle=\color{mauve},
  breaklines=true,
  breakatwhitespace=true
  tabsize=3
}

@@

	Consider the problem
	\[SEND + MORE = MONEY\]
	where every character is a variable and the position in the word describes the weight.
	\begin{equation}
	\label{eq:linearConstraint}
	\begin{split}
		& S \cdot 10^3 + E \cdot 10^2 + N \cdot 10^1 + D \cdot 10^0 + \\ 
		& M \cdot 10^3 + O \cdot 10^2 + R \cdot 10^1 + E \cdot 10^0 = \\ 
		& M \cdot 10^4 + O \cdot 10^3 + N \cdot 10^2 + E \cdot 10^1 + Y \cdot 10^0
	\end{split}
	\end{equation}

	To narrow the solution space a little bit down I state additional constraints:
	\begin{enumerate}
	\item All variables must have different values.
		\begin{equation}
		\begin{split}
			S \ne E \land S \ne N \land & S \ne D \land \dots \land \\
			E \ne N \land & E \ne D \land \dots \land \\
			& N \ne D \land \dots
		\end{split}
		\end{equation}	
	\item There must not be a trailing zero. Combined with the so-called all-different constraint \cite{allDifferent} from above this results in
		\begin{equation}
			S \ne 0 \land M \ne 0
		\end{equation}
	\end{enumerate}

@@

	\subsection{Constraint Types}
	As already mentioned before constraint programming is the generalization of linear programming, therefore it is at least possible do define linear constraints like \ref{eq:linearConstraint}. Which statements are possible to make varies than very much from the programming language in which the problem is implemented and the used toolset. Typical constraints would be
	\begin{itemize}
		\item Equalities: $x = y + 2$, $x^2 + y^2 = 1$, \dots
		\item Inequalities: $x \ne y$, \dots
		\item Relations: $x < 2 y$, $e^x < 4$, \dots
	\end{itemize}
	In Gecode for example it is very easy to create for example a nonlinear constraint based on the integer variables a, b, c and d \cite[p. 120]{programmingGecode}:
	\begin{lstlisting}
rel(home, a+b*(c+d) == 0);
	\end{lstlisting}
	This single statement in C++ implements the constraint
	\begin{equation}
		a + b \cdot (c + d) == 0
	\end{equation}
	what is a quite handy way to do it.

	Hence there are boolean variables available it is also possible to use logical operators to combine other constraint types to logical expressions. To show the usage of this on an example I will give the implementation for the placement of two squares such that they do not overlap. This problem can be expressed mathematically by \cite[p. 101]{programmingGecode}
	\begin{equation}
	\begin{split}
		x_1 + d_1 \le x_2 & \lor x_2 + d_2 \le x_1 \lor \\
		y_1 + d_1 \le y_2 & \lor y_2 + d_2 \le y_1
	\end{split}
	\end{equation}
	and in Gecode by
	\begin{lstlisting}
rel(home,	(x1 + d1 <= x2) || 
		(x2 + d2 <= x1) || 
		(y1 + d1 <= y2) || 
		(y2 + d2 <= y1));
	\end{lstlisting}
	which is again quite obvious to a reader with some experience in a C-like programming language.

@@

	\begin{thebibliography}{1}
		\bibitem{handbookCP}
		F.~Rossi, P.van~Beek and T.Walsh, \emph{Handbook of Constraint Programming}, Amsterdam, The Netherlands: Elsevier, 2006, ISBN: 978-0-444-52726-4
		\bibitem{allDifferent}
		W.~van~Hoeve, \emph{The Alldifferent Constraint: A Survey}, http://www.andrew.cmu.edu/user/vanhoeve/papers/alldiff.pdf
		\bibitem{trendsInCP}
		F.~Benhamou, N.~Jussien, B.A.O'Sullivan, \emph{Trends in Constraint Programming}, Wiley-ISTE, 2007, ISBN: 978-1-905209-97-2
		\bibitem{gecode}
		http://www.gecode.org/
		\bibitem{programmingGecode}
		C.~Schulte, G.~Tack and M.Z.Lagerkvist, \emph{Modeling and Programming with Gecode}, http://www.gecode.org/doc-latest/MPG.pdf
	\end{thebibliography}
\end{document}
